---
title: "HW10"
author: "Avinash Ramu"
date: "November 28, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q21.3
A) Fit a varying intercept varying slope model to the radon data. Get point estimates(median) and 50% intervals for the finite population and superpopulation standard-deviations using this data.

I will use Andy's code from earlier to setup the model.

```{r q1_1, include=FALSE}
##Use Andy's tutorial code to setup the radon data.
# Setting up the data (from Gelman & Hill's website)
library(rjags); library(R2jags)
srrs2 <- read.table ("srrs2.dat", header=T, sep=",")
mn <- srrs2$state=="MN"
srrs2_mn <- srrs2[mn, ]
radon <- srrs2$activity[mn]
log.radon <- log (ifelse (radon==0, .1, radon))
floor <- srrs2$floor[mn]       # 0 for basement, 1 for first floor
n <- length(radon)
y <- log.radon
x <- floor

# get county index variable
county.name <- as.vector(srrs2$county[mn])
uniq <- unique(county.name)
J <- length(uniq)
county <- rep (NA, J)
for (i in 1:J){
  county[county.name==uniq[i]] <- i
}

srrs2.fips <- srrs2$stfips*1000 + srrs2$cntyfips
cty <- read.table ("cty.dat", header=T, sep=",")
usa.fips <- 1000*cty[,"stfips"] + cty[,"ctfips"]
usa.rows <- match (unique(srrs2.fips[mn]), usa.fips)
uranium <- cty[usa.rows,"Uppm"]
u <- log (uranium)


the.model <- "model { 
  for (i in 1:n){
    y[i] ~ dnorm (y.hat[i], tau.y)
    y.hat[i] <- a[county[i]] + b[county[i]]*x[i]
    e.y[i] = y[i] - y.hat[i]
  }
  s.y <- sd(e.y[])
  tau.y <- pow(sigma.y, -2)
  sigma.y ~ dunif (0, 100)
  for (j in 1:J){
    a[j] ~ dnorm (a.hat[j], tau.a)
    b[j] ~ dnorm (b.hat[j], tau.b)
    a.hat[j] <- mu.a
    b.hat[j] <- mu.b
    e.a[j] <- a[j] - a.hat[j]
    e.b[j] <- b[j] - b.hat[j]
  }
  s.a <- sd(e.a[])
  s.b <- sd(e.b[])
  mu.a ~ dnorm (0, .0001)
  mu.b ~ dnorm (0, .0001)
  tau.a <- pow(sigma.a, -2)
  tau.b <- pow(sigma.b, -2)
  sigma.a ~ dunif (0, 100)
  sigma.b ~ dunif (0, 100)
}"

# Defining the data you will pass into the model -- you *already know* these values
radon.data <- list ("n", "J", "x", "y", "county", "u")

# Defining the initial values that your model's parameters (values you *don't* already know)
radon.inits <- function (){list(a=rnorm(J), 
                                b=rnorm(J), 
                                g.0=rnorm(1), 
                                g.1=rnorm(1),
                                sigma.y=runif(1), 
                                sigma.a=runif(1))}

# Defining which parameters of your model you want JAGS to return to you 
# In the book (page 366), they are missing "g.0" and "g.1"
radon.parameters <- c ("a", "b", "sigma.y", "sigma.a", "sigma.b", "s.y", "s.a", "s.b", "g.0", "g.1")

# Now, we can actually run the model with the jags() function
radon.3 <- jags(data=radon.data, 
                inits=radon.inits, 
                parameters.to.save=radon.parameters, 
                model.file=textConnection(the.model), # Note the textConnection() function
                n.chains=3, 
                n.iter=500, 
                DIC=F)
```

```{r q1_2}
plot(radon.3)
radon.3
attach.jags(radon.3)

#Toss out the burn-in values
sigma.a <- sigma.a[251:750]
sigma.b <- sigma.b[251:750]
sigma.y <- sigma.y[251:750]
s.a <- s.a[251:750]
s.b <- s.b[251:750]
s.y <- s.y[251:750]

sigma.a_mean <- mean(sigma.a)
sigma.b_mean <- mean(sigma.b)
sigma.y_mean <- mean(sigma.y)
s.a_mean <- mean(s.a)
s.b_mean <- mean(s.b)
s.y_mean <- mean(s.y)

#sigma.a and s.a
print(c(sigma.a_mean, quantile(sigma.a, c(0.25, 0.75))))
print(c(s.a_mean, quantile(s.a, c(0.25, 0.75))))

#sigma.b and s.b
print(c(sigma.b_mean, quantile(sigma.b, c(0.25, 0.75))))
print(c(s.b_mean, quantile(s.b, c(0.25, 0.75))))

#sigma.y and s.y
print(c(sigma.y_mean, quantile(sigma.y, c(0.25, 0.75))))
print(c(s.y_mean, quantile(s.y, c(0.25, 0.75))))
```

```{r q1_12, include=FALSE}
sample_counties <- sample(srrs2_mn$county, 10)
srrs2 <- srrs2[srrs2$county %in% sample_counties, ]
mn <- srrs2$state=="MN"
srrs2_mn <- srrs2[mn, ]
radon <- srrs2$activity[mn]
log.radon <- log (ifelse (radon==0, .1, radon))
floor <- srrs2$floor[mn]       # 0 for basement, 1 for first floor
n <- length(radon)
y <- log.radon
x <- floor

# get county index variable
county.name <- as.vector(srrs2$county[mn])
uniq <- unique(county.name)
J <- length(uniq)
county <- rep (NA, J)
for (i in 1:J){
  county[county.name==uniq[i]] <- i
}

srrs2.fips <- srrs2$stfips*1000 + srrs2$cntyfips
cty <- read.table ("cty.dat", header=T, sep=",")
usa.fips <- 1000*cty[,"stfips"] + cty[,"ctfips"]
usa.rows <- match (unique(srrs2.fips[mn]), usa.fips)
uranium <- cty[usa.rows,"Uppm"]
u <- log (uranium)


the.model <- "model { 
  for (i in 1:n){
    y[i] ~ dnorm (y.hat[i], tau.y)
    y.hat[i] <- a[county[i]] + b[county[i]]*x[i]
    e.y[i] = y[i] - y.hat[i]
  }
  s.y <- sd(e.y[])
  tau.y <- pow(sigma.y, -2)
  sigma.y ~ dunif (0, 100)
  for (j in 1:J){
    a[j] ~ dnorm (a.hat[j], tau.a)
    b[j] ~ dnorm (b.hat[j], tau.b)
    a.hat[j] <- mu.a
    b.hat[j] <- mu.b
    e.a[j] <- a[j] - a.hat[j]
    e.b[j] <- b[j] - b.hat[j]
  }
  s.a <- sd(e.a[])
  s.b <- sd(e.b[])
  mu.a ~ dnorm (0, .0001)
  mu.b ~ dnorm (0, .0001)
  tau.a <- pow(sigma.a, -2)
  tau.b <- pow(sigma.b, -2)
  sigma.a ~ dunif (0, 100)
  sigma.b ~ dunif (0, 100)
}"

# Defining the data you will pass into the model -- you *already know* these values
radon.data <- list ("n", "J", "x", "y", "county", "u")

# Defining the initial values that your model's parameters (values you *don't* already know)
radon.inits <- function (){list(a=rnorm(J), 
                                b=rnorm(J), 
                                g.0=rnorm(1), 
                                g.1=rnorm(1),
                                sigma.y=runif(1), 
                                sigma.a=runif(1))}

# Defining which parameters of your model you want JAGS to return to you 
# In the book (page 366), they are missing "g.0" and "g.1"
radon.parameters <- c ("a", "b", "sigma.y", "sigma.a", "sigma.b", "s.y", "s.a", "s.b", "g.0", "g.1")

# Now, we can actually run the model with the jags() function
radon.3 <- jags(data=radon.data, 
                inits=radon.inits, 
                parameters.to.save=radon.parameters, 
                model.file=textConnection(the.model), # Note the textConnection() function
                n.chains=3, 
                n.iter=500, 
                DIC=F)
```

```{r q1_22}
attach.jags(radon.3)

#Toss out the burn-in values
sigma.a <- sigma.a[251:750]
sigma.b <- sigma.b[251:750]
sigma.y <- sigma.y[251:750]
s.a <- s.a[251:750]
s.b <- s.b[251:750]
s.y <- s.y[251:750]

sigma.a_mean <- mean(sigma.a, na.rm = T)
sigma.b_mean <- mean(sigma.b, na.rm = T)
sigma.y_mean <- mean(sigma.y, na.rm = T)
s.a_mean <- mean(s.a, na.rm = T)
s.b_mean <- mean(s.b, na.rm = T)
s.y_mean <- mean(s.y, na.rm = T)

#sigma.a and s.a
print(c(sigma.a_mean, quantile(sigma.a, c(0.25, 0.75), na.rm = T)))
print(c(s.a_mean, quantile(s.a, c(0.25, 0.75), na.rm = T)))

#sigma.b and s.b
print(c(sigma.b_mean, quantile(sigma.b, c(0.25, 0.75), na.rm = T)))
print(c(s.b_mean, quantile(s.b, c(0.25, 0.75), na.rm = T)))

#sigma.y and s.y
print(c(sigma.y_mean, quantile(sigma.y, c(0.25, 0.75), na.rm = T)))
print(c(s.y_mean, quantile(s.y, c(0.25, 0.75), na.rm = T)))
```

## 21.4
Fit in Bugs a varying-intercept model to the radon data with log
uranium as a group-level predictor. You will compare inferences for the superpopulation
contrast (that is, the slope for log uranium in the county-level model)
and the corresponding finite-population contrast (that is, the coefficient of log
uranium for the intercepts for the particular counties in the data). You will need
to postprocess the simulations in R in order to get simulations for the finitepopulation
contrast.

I used the code from Andy as a reference for the varying intercept model.

```{r q2_1, include=FALSE}
##Use Andy's tutorial code to setup the radon data.
# Setting up the data (from Gelman & Hill's website)
library(rjags); library(R2jags)
srrs2 <- read.table ("srrs2.dat", header=T, sep=",")
mn <- srrs2$state=="MN"
srrs2_mn <- srrs2[mn, ]
radon <- srrs2$activity[mn]
log.radon <- log (ifelse (radon==0, .1, radon))
floor <- srrs2$floor[mn]       # 0 for basement, 1 for first floor
n <- length(radon)
y <- log.radon
x <- floor

# get county index variable
county.name <- as.vector(srrs2$county[mn])
uniq <- unique(county.name)
J <- length(uniq)
county <- rep (NA, J)
for (i in 1:J){
  county[county.name==uniq[i]] <- i
}

srrs2.fips <- srrs2$stfips*1000 + srrs2$cntyfips
cty <- read.table ("cty.dat", header=T, sep=",")
usa.fips <- 1000*cty[,"stfips"] + cty[,"ctfips"]
usa.rows <- match (unique(srrs2.fips[mn]), usa.fips)
uranium <- cty[usa.rows,"Uppm"]
u <- log (uranium)

the.model <- "model{
  for (i in 1:n){
    y[i] ~ dnorm (y.hat[i], tau.y)
    y.hat[i] <- a[county[i]] + b*x[i]
  }
  b ~ dnorm (0, .0001)
  tau.y <- pow(sigma.y, -2)
  sigma.y ~ dunif (0, 100)
  
  for (j in 1:J){
    a[j] ~ dnorm (a.hat[j], tau.a)
    a.hat[j] <- g.0 + g.1*u[j]
  }
  g.0 ~ dnorm (0, .0001)
  g.1 ~ dnorm (0, .0001)
  tau.a <- pow(sigma.a, -2)
  sigma.a ~ dunif (0, 100)
}"

# Defining the data you will pass into the model -- you *already know* these values
radon.data <- list ("n", "J", "x", "y", "county", "u")

# Defining the initial values that your model's parameters (values you *don't* already know)
radon.inits <- function (){list(a=rnorm(J), 
                                b=rnorm(1), 
                                g.0=rnorm(1), 
                                g.1=rnorm(1),
                                sigma.y=runif(1), 
                                sigma.a=runif(1))}

# Defining which parameters of your model you want JAGS to return to you 
# In the book (page 366), they are missing "g.0" and "g.1"
radon.parameters <- c ("a", "b", "sigma.y", "sigma.a", "g.0", "g.1")

# Now, we can actually run the model with the jags() function
radon.3 <- jags(data=radon.data, 
                inits=radon.inits, 
                parameters.to.save=radon.parameters, 
                model.file=textConnection(the.model), # Note the textConnection() function
                n.chains=3, 
                n.iter=500, 
                DIC=F)

radon.3
```


Get the contrasts
```{r q2_2}
n.sims <- 750
attach.jags(radon.3)
mean(g.1)
quantile(g.1, c(.025,.975), na.rm = T)
finite.slope <- rep (NA, 500)
for (s in 251:n.sims){
  finite.pop <- lm (a[s,] ~ u)
  finite.slope[s] <- coef(finite.pop)["u"]
}
mean(finite.slope, na.rm = T)
quantile (finite.slope, c(.025,.975), na.rm = T)
```
The superpopulation and finite-population contrasts are very similar in value.

##Part B - only look at the first three counties
I don't show the model fitting code similar to previous downsampling

```{r q2_3, include = FALSE}
##Use Andy's tutorial code to setup the radon data.
# Setting up the data (from Gelman & Hill's website)
library(rjags); library(R2jags)
srrs2 <- read.table ("srrs2.dat", header=T, sep=",")
mn <- srrs2$state=="MN"
srrs2_mn <- srrs2[mn, ]
sample_counties <- sample(srrs2_mn$county, 3)
srrs2 <- srrs2[srrs2$county %in% sample_counties, ]
mn <- srrs2$state=="MN"
srrs2_mn <- srrs2[mn, ]
radon <- srrs2$activity[mn]
log.radon <- log (ifelse (radon==0, .1, radon))
floor <- srrs2$floor[mn]       # 0 for basement, 1 for first floor
n <- length(radon)
y <- log.radon
x <- floor

# get county index variable
county.name <- as.vector(srrs2$county[mn])
uniq <- unique(county.name)
J <- length(uniq)
county <- rep (NA, J)
for (i in 1:J){
  county[county.name==uniq[i]] <- i
}

srrs2.fips <- srrs2$stfips*1000 + srrs2$cntyfips
cty <- read.table ("cty.dat", header=T, sep=",")
usa.fips <- 1000*cty[,"stfips"] + cty[,"ctfips"]
usa.rows <- match (unique(srrs2.fips[mn]), usa.fips)
uranium <- cty[usa.rows,"Uppm"]
u <- log (uranium)
print(u)

the.model <- "model{
  for (i in 1:n){
    y[i] ~ dnorm (y.hat[i], tau.y)
    y.hat[i] <- a[county[i]] + b*x[i]
  }
  b ~ dnorm (0, .0001)
  tau.y <- pow(sigma.y, -2)
  sigma.y ~ dunif (0, 100)
  
  for (j in 1:J){
    a[j] ~ dnorm (a.hat[j], tau.a)
    a.hat[j] <- g.0 + g.1*u[j]
  }
  g.0 ~ dnorm (0, .0001)
  g.1 ~ dnorm (0, .0001)
  tau.a <- pow(sigma.a, -2)
  sigma.a ~ dunif (0, 100)
}"

# Defining the data you will pass into the model -- you *already know* these values
radon.data <- list ("n", "J", "x", "y", "county", "u")

# Defining the initial values that your model's parameters (values you *don't* already know)
radon.inits <- function (){list(a=rnorm(J), 
                                b=rnorm(1), 
                                g.0=rnorm(1), 
                                g.1=rnorm(1),
                                sigma.y=runif(1), 
                                sigma.a=runif(1))}

# Defining which parameters of your model you want JAGS to return to you 
# In the book (page 366), they are missing "g.0" and "g.1"
radon.parameters <- c ("a", "b", "sigma.y", "sigma.a", "g.0", "g.1")

# Now, we can actually run the model with the jags() function
radon.3 <- jags(data=radon.data, 
                inits=radon.inits, 
                parameters.to.save=radon.parameters, 
                model.file=textConnection(the.model), # Note the textConnection() function
                n.chains=3, 
                n.iter=500, 
                DIC=F)

radon.3
```

Get the contrasts for the model using only three counties.
Note I sampled three random counties instead of picking the first three,
the same inferences should hold.

```{r q2_4}
n.sims <- 750
attach.jags(radon.3)
mean(g.1)
quantile(g.1, c(.025,.975), na.rm = T)
finite.slope <- rep (NA, 500)
for (s in 251:n.sims){
  finite.pop <- lm (a[s,] ~ u)
  finite.slope[s] <- coef(finite.pop)["u"]
}
mean(finite.slope, na.rm = T)
quantile (finite.slope, c(.025,.975), na.rm = T)
```

The superpopulation and finite-population contrasts are vastly different in this case.