---
title: "HW10"
author: "Avinash Ramu"
date: "November 28, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Q21.1

1.
If more houses are added per county, the parameter estimates will likely be the same but the
finite population standard deviations will go down i.e the uncertainity in the data drops.
The superpopulation standard deviation will remain the same since we have the same number
of counties. Another way of stating this is that the uncertainity in individual alpha_j and mu_j goes down but the collective uncertainity remains the same.

2.
If 4 times as many counties are added, the parameter estimates will likely remain the same
but the superpopulation standard deviation will decrease, i.e uncertainity about mu_alpha and sigma_alpha goes down. We are able to better fit the distribution for the alpha_j . This is because we have a larger sample size for alpha_j
and we understand this distribution better with more counties. The within county inferences
will likely remain the same.

3. 
If we have 4 times as many houses and 4 times as many counties, both the finite population
standard deviation and the superpopulation standard deviations should decrease. This is because
we have more samples at both levels. Assuming that our initial samples were a good representation of the underlying population the parameter estimates should not change significantly. Another way of stating this is that the uncertainity of individual alpha_j and mu_j will go down, also uncertainity about the distribution of mu_j s will go down.

## Q21.3
A) Fit a varying intercept varying slope model to the radon data. Get point estimates(median) and 50% intervals for the finite population and superpopulation standard-deviations using this data.

I will use Andy's code from earlier to setup the model.

```{r q1_1, include=FALSE}
##Use Andy's tutorial code to setup the radon data.
# Setting up the data (from Gelman & Hill's website)
library(rjags); library(R2jags)
srrs2 <- read.table ("srrs2.dat", header=T, sep=",")
mn <- srrs2$state=="MN"
srrs2_mn <- srrs2[mn, ]
radon <- srrs2$activity[mn]
log.radon <- log (ifelse (radon==0, .1, radon))
floor <- srrs2$floor[mn]       # 0 for basement, 1 for first floor
n <- length(radon)
y <- log.radon
x <- floor

# get county index variable
county.name <- as.vector(srrs2$county[mn])
uniq <- unique(county.name)
J <- length(uniq)
county <- rep (NA, J)
for (i in 1:J){
  county[county.name==uniq[i]] <- i
}

srrs2.fips <- srrs2$stfips*1000 + srrs2$cntyfips
cty <- read.table ("cty.dat", header=T, sep=",")
usa.fips <- 1000*cty[,"stfips"] + cty[,"ctfips"]
usa.rows <- match (unique(srrs2.fips[mn]), usa.fips)
uranium <- cty[usa.rows,"Uppm"]
u <- log (uranium)


the.model <- "model { 
  for (i in 1:n){
    y[i] ~ dnorm (y.hat[i], tau.y)
    y.hat[i] <- a[county[i]] + b[county[i]]*x[i]
    e.y[i] = y[i] - y.hat[i]
  }
  s.y <- sd(e.y[])
  tau.y <- pow(sigma.y, -2)
  sigma.y ~ dunif (0, 100)
  for (j in 1:J){
    a[j] ~ dnorm (a.hat[j], tau.a)
    b[j] ~ dnorm (b.hat[j], tau.b)
    a.hat[j] <- mu.a
    b.hat[j] <- mu.b
    e.a[j] <- a[j] - a.hat[j]
    e.b[j] <- b[j] - b.hat[j]
  }
  s.a <- sd(e.a[])
  s.b <- sd(e.b[])
  mu.a ~ dnorm (0, .0001)
  mu.b ~ dnorm (0, .0001)
  tau.a <- pow(sigma.a, -2)
  tau.b <- pow(sigma.b, -2)
  sigma.a ~ dunif (0, 100)
  sigma.b ~ dunif (0, 100)
}"

# Defining the data you will pass into the model -- you *already know* these values
radon.data <- list ("n", "J", "x", "y", "county", "u")

# Defining the initial values that your model's parameters (values you *don't* already know)
radon.inits <- function (){list(a=rnorm(J), 
                                b=rnorm(J), 
                                g.0=rnorm(1), 
                                g.1=rnorm(1),
                                sigma.y=runif(1), 
                                sigma.a=runif(1))}

# Defining which parameters of your model you want JAGS to return to you 
# In the book (page 366), they are missing "g.0" and "g.1"
radon.parameters <- c ("a", "b", "sigma.y", "sigma.a", "sigma.b", "s.y", "s.a", "s.b", "g.0", "g.1")

# Now, we can actually run the model with the jags() function
radon.3 <- jags(data=radon.data, 
                inits=radon.inits, 
                parameters.to.save=radon.parameters, 
                model.file=textConnection(the.model), # Note the textConnection() function
                n.chains=3, 
                n.iter=500, 
                DIC=F)
```

```{r q1_2}
plot(radon.3)
attach.jags(radon.3, overwrite = T)

#Toss out the burn-in values
sigma.a <- sigma.a[251:750]
sigma.b <- sigma.b[251:750]
sigma.y <- sigma.y[251:750]
s.a <- s.a[251:750]
s.b <- s.b[251:750]
s.y <- s.y[251:750]

sigma.a_mean <- mean(sigma.a)
sigma.b_mean <- mean(sigma.b)
sigma.y_mean <- mean(sigma.y)
s.a_mean <- mean(s.a)
s.b_mean <- mean(s.b)
s.y_mean <- mean(s.y)

#sigma.a and s.a
print(c(sigma.a_mean, quantile(sigma.a, c(0.25, 0.75))))
print(c(s.a_mean, quantile(s.a, c(0.25, 0.75))))

#sigma.b and s.b
print(c(sigma.b_mean, quantile(sigma.b, c(0.25, 0.75))))
print(c(s.b_mean, quantile(s.b, c(0.25, 0.75))))

#sigma.y and s.y
print(c(sigma.y_mean, quantile(sigma.y, c(0.25, 0.75))))
print(c(s.y_mean, quantile(s.y, c(0.25, 0.75))))
```

```{r q1_12, include=FALSE}
sample_counties <- sample(srrs2_mn$county, 10)
srrs2 <- srrs2[srrs2$county %in% sample_counties, ]
mn <- srrs2$state=="MN"
srrs2_mn <- srrs2[mn, ]
radon <- srrs2$activity[mn]
log.radon <- log (ifelse (radon==0, .1, radon))
floor <- srrs2$floor[mn]       # 0 for basement, 1 for first floor
n <- length(radon)
y <- log.radon
x <- floor

# get county index variable
county.name <- as.vector(srrs2$county[mn])
uniq <- unique(county.name)
J <- length(uniq)
county <- rep (NA, J)
for (i in 1:J){
  county[county.name==uniq[i]] <- i
}

srrs2.fips <- srrs2$stfips*1000 + srrs2$cntyfips
cty <- read.table ("cty.dat", header=T, sep=",")
usa.fips <- 1000*cty[,"stfips"] + cty[,"ctfips"]
usa.rows <- match (unique(srrs2.fips[mn]), usa.fips)
uranium <- cty[usa.rows,"Uppm"]
u <- log (uranium)


the.model <- "model { 
  for (i in 1:n){
    y[i] ~ dnorm (y.hat[i], tau.y)
    y.hat[i] <- a[county[i]] + b[county[i]]*x[i]
    e.y[i] = y[i] - y.hat[i]
  }
  s.y <- sd(e.y[])
  tau.y <- pow(sigma.y, -2)
  sigma.y ~ dunif (0, 100)
  for (j in 1:J){
    a[j] ~ dnorm (a.hat[j], tau.a)
    b[j] ~ dnorm (b.hat[j], tau.b)
    a.hat[j] <- mu.a
    b.hat[j] <- mu.b
    e.a[j] <- a[j] - a.hat[j]
    e.b[j] <- b[j] - b.hat[j]
  }
  s.a <- sd(e.a[])
  s.b <- sd(e.b[])
  mu.a ~ dnorm (0, .0001)
  mu.b ~ dnorm (0, .0001)
  tau.a <- pow(sigma.a, -2)
  tau.b <- pow(sigma.b, -2)
  sigma.a ~ dunif (0, 100)
  sigma.b ~ dunif (0, 100)
}"

# Defining the data you will pass into the model -- you *already know* these values
radon.data <- list ("n", "J", "x", "y", "county", "u")

# Defining the initial values that your model's parameters (values you *don't* already know)
radon.inits <- function (){list(a=rnorm(J), 
                                b=rnorm(J), 
                                g.0=rnorm(1), 
                                g.1=rnorm(1),
                                sigma.y=runif(1), 
                                sigma.a=runif(1))}

# Defining which parameters of your model you want JAGS to return to you 
# In the book (page 366), they are missing "g.0" and "g.1"
radon.parameters <- c ("a", "b", "sigma.y", "sigma.a", "sigma.b", "s.y", "s.a", "s.b", "g.0", "g.1")

# Now, we can actually run the model with the jags() function
radon.3 <- jags(data=radon.data, 
                inits=radon.inits, 
                parameters.to.save=radon.parameters, 
                model.file=textConnection(the.model), # Note the textConnection() function
                n.chains=3, 
                n.iter=500, 
                DIC=F)
```

```{r q1_22}
attach.jags(radon.3, overwrite = T)

#Toss out the burn-in values
sigma.a <- sigma.a[251:750]
sigma.b <- sigma.b[251:750]
sigma.y <- sigma.y[251:750]
s.a <- s.a[251:750]
s.b <- s.b[251:750]
s.y <- s.y[251:750]

sigma.a_mean <- mean(sigma.a, na.rm = T)
sigma.b_mean <- mean(sigma.b, na.rm = T)
sigma.y_mean <- mean(sigma.y, na.rm = T)
s.a_mean <- mean(s.a, na.rm = T)
s.b_mean <- mean(s.b, na.rm = T)
s.y_mean <- mean(s.y, na.rm = T)

#sigma.a and s.a
print(c(sigma.a_mean, quantile(sigma.a, c(0.25, 0.75), na.rm = T)))
print(c(s.a_mean, quantile(s.a, c(0.25, 0.75), na.rm = T)))

#sigma.b and s.b
print(c(sigma.b_mean, quantile(sigma.b, c(0.25, 0.75), na.rm = T)))
print(c(s.b_mean, quantile(s.b, c(0.25, 0.75), na.rm = T)))

#sigma.y and s.y
print(c(sigma.y_mean, quantile(sigma.y, c(0.25, 0.75), na.rm = T)))
print(c(s.y_mean, quantile(s.y, c(0.25, 0.75), na.rm = T)))
```

## 21.4
Fit in Bugs a varying-intercept model to the radon data with log
uranium as a group-level predictor. You will compare inferences for the superpopulation
contrast (that is, the slope for log uranium in the county-level model)
and the corresponding finite-population contrast (that is, the coefficient of log
uranium for the intercepts for the particular counties in the data). You will need
to postprocess the simulations in R in order to get simulations for the finitepopulation
contrast.

I used the code from Andy as a reference for the varying intercept model.

```{r q2_1, include=FALSE}
##Use Andy's tutorial code to setup the radon data.
# Setting up the data (from Gelman & Hill's website)
library(rjags); library(R2jags)
srrs2 <- read.table ("srrs2.dat", header=T, sep=",")
mn <- srrs2$state=="MN"
srrs2_mn <- srrs2[mn, ]
radon <- srrs2$activity[mn]
log.radon <- log (ifelse (radon==0, .1, radon))
floor <- srrs2$floor[mn]       # 0 for basement, 1 for first floor
n <- length(radon)
y <- log.radon
x <- floor

# get county index variable
county.name <- as.vector(srrs2$county[mn])
uniq <- unique(county.name)
J <- length(uniq)
county <- rep (NA, J)
for (i in 1:J){
  county[county.name==uniq[i]] <- i
}

srrs2.fips <- srrs2$stfips*1000 + srrs2$cntyfips
cty <- read.table ("cty.dat", header=T, sep=",")
usa.fips <- 1000*cty[,"stfips"] + cty[,"ctfips"]
usa.rows <- match (unique(srrs2.fips[mn]), usa.fips)
uranium <- cty[usa.rows,"Uppm"]
u <- log (uranium)

the.model <- "model{
  for (i in 1:n){
    y[i] ~ dnorm (y.hat[i], tau.y)
    y.hat[i] <- a[county[i]] + b*x[i]
    e.y[i] <- y[i] - y.hat[i]
  }
  s.y <- sd(e.y[])
  b ~ dnorm (0, .0001)
  tau.y <- pow(sigma.y, -2)
  sigma.y ~ dunif (0, 100)
  for (j in 1:J){
    a[j] ~ dnorm (a.hat[j], tau.a)
    a.hat[j] <- g.0 + g.1*u[j]
    e.a[j] <- a[j] - a.hat[j]
  }
  s.a <- sd(e.a[])
  g.0 ~ dnorm (0, .0001)
  g.1 ~ dnorm (0, .0001)
  tau.a <- pow(sigma.a, -2)
  sigma.a ~ dunif (0, 100)
}"

# Defining the data you will pass into the model -- you *already know* these values
radon.data <- list ("n", "J", "x", "y", "county", "u")

# Defining the initial values that your model's parameters (values you *don't* already know)
radon.inits <- function (){list(a=rnorm(J), 
                                b=rnorm(1), 
                                g.0=rnorm(1), 
                                g.1=rnorm(1),
                                sigma.y=runif(1), 
                                sigma.a=runif(1))}

# Defining which parameters of your model you want JAGS to return to you 
radon.parameters <- c ("a", "b", "sigma.y", "sigma.a", "s.a", "s.y", "g.0", "g.1")

# Now, we can actually run the model with the jags() function
radon.3 <- jags(data=radon.data, 
                inits=radon.inits, 
                parameters.to.save=radon.parameters, 
                model.file=textConnection(the.model), # Note the textConnection() function
                n.chains=3, 
                n.iter=500, 
                DIC=F)

radon.3
radon.3_all_fixed_intercept <- radon.3
```


Get the contrasts
```{r q2_2}
n.sims <- 750
attach.jags(radon.3, overwrite = T)
mean(g.1)
quantile(g.1, c(.025,.975), na.rm = T)
finite.slope <- rep (NA, 500)
for (s in 251:n.sims){
  finite.pop <- lm (a[s,] ~ u)
  finite.slope[s] <- coef(finite.pop)["u"]
}
mean(finite.slope, na.rm = T)
quantile (finite.slope, c(.025,.975), na.rm = T)
```
The superpopulation and finite-population contrasts are very similar in value.

##Part B - only look at the first three counties
I don't show the model fitting code, this code is similar to previous downsampling examples in this homework.

```{r q2_3, include = FALSE}
##Use Andy's tutorial code to setup the radon data.
# Setting up the data (from Gelman & Hill's website)
library(rjags); library(R2jags)
srrs2 <- read.table ("srrs2.dat", header=T, sep=",")
mn <- srrs2$state=="MN"
srrs2_mn <- srrs2[mn, ]
sample_counties <- sample(srrs2_mn$county, 3)
srrs2 <- srrs2[srrs2$county %in% sample_counties, ]
mn <- srrs2$state=="MN"
srrs2_mn <- srrs2[mn, ]
radon <- srrs2$activity[mn]
log.radon <- log (ifelse (radon==0, .1, radon))
floor <- srrs2$floor[mn]       # 0 for basement, 1 for first floor
n <- length(radon)
y <- log.radon
x <- floor

# get county index variable
county.name <- as.vector(srrs2$county[mn])
uniq <- unique(county.name)
J <- length(uniq)
county <- rep (NA, J)
for (i in 1:J){
  county[county.name==uniq[i]] <- i
}

srrs2.fips <- srrs2$stfips*1000 + srrs2$cntyfips
cty <- read.table ("cty.dat", header=T, sep=",")
usa.fips <- 1000*cty[,"stfips"] + cty[,"ctfips"]
usa.rows <- match (unique(srrs2.fips[mn]), usa.fips)
uranium <- cty[usa.rows,"Uppm"]
u <- log (uranium)
print(u)

the.model <- "model{
  for (i in 1:n){
    y[i] ~ dnorm (y.hat[i], tau.y)
    y.hat[i] <- a[county[i]] + b*x[i]
  }
  b ~ dnorm (0, .0001)
  tau.y <- pow(sigma.y, -2)
  sigma.y ~ dunif (0, 100)
  
  for (j in 1:J){
    a[j] ~ dnorm (a.hat[j], tau.a)
    a.hat[j] <- g.0 + g.1*u[j]
  }
  g.0 ~ dnorm (0, .0001)
  g.1 ~ dnorm (0, .0001)
  tau.a <- pow(sigma.a, -2)
  sigma.a ~ dunif (0, 100)
}"

# Defining the data you will pass into the model -- you *already know* these values
radon.data <- list ("n", "J", "x", "y", "county", "u")

# Defining the initial values that your model's parameters (values you *don't* already know)
radon.inits <- function (){list(a=rnorm(J), 
                                b=rnorm(1), 
                                g.0=rnorm(1), 
                                g.1=rnorm(1),
                                sigma.y=runif(1), 
                                sigma.a=runif(1))}

# Defining which parameters of your model you want JAGS to return to you 
# In the book (page 366), they are missing "g.0" and "g.1"
radon.parameters <- c ("a", "b", "sigma.y", "sigma.a", "g.0", "g.1")

# Now, we can actually run the model with the jags() function
radon.3 <- jags(data=radon.data, 
                inits=radon.inits, 
                parameters.to.save=radon.parameters, 
                model.file=textConnection(the.model), # Note the textConnection() function
                n.chains=3, 
                n.iter=500, 
                DIC=F)

radon.3
```

Get the contrasts for the model using only three counties.
Note -  I sampled three random counties instead of picking the first three,
the same inferences should hold.

```{r q2_4}
n.sims <- 750
attach.jags(radon.3, overwrite = T)
mean(g.1)
quantile(g.1, c(.025,.975), na.rm = T)
finite.slope <- rep (NA, 500)
for (s in 251:n.sims){
  finite.pop <- lm (a[s,] ~ u)
  finite.slope[s] <- coef(finite.pop)["u"]
}
mean(finite.slope, na.rm = T)
quantile (finite.slope, c(.025,.975), na.rm = T)
```

The superpopulation and finite-population contrasts are vastly different in this case.


## Q22.1

Take a varying-intercept model from one of the exercises in Part 2 of this book
and construct the corresponding ANOVA plot as in Section 22.3.

I will use the varying intercept radon model for this question.
Fit the model using the same framework described above.

```{r q3_1}
rm(s.a)
rm(s.y)
attach.jags(radon.3_all_fixed_intercept, overwrite = T)
combined <- data.frame(index = 251:750, s.a = s.a[251:750], s.y = s.y[251:750])
require(reshape2)
require(ggplot2)
combined_m <- melt(combined, id.vars = c("index"))
ggplot(combined_m) + geom_boxplot(aes(x = variable, y = value))
```