---
title: "HW11"
author: "Avinash Ramu"
date: "December 6, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 24.1
Download the data in the folder dogs and fit some other models, for example
using as a predictor the result from the previous trial, or the previous two trials,
rather than the total number of shocks and avoidances.

(a) Fit this model, as usual building up from simpler versions (first a singlelevel
model, then varying intercepts, then varying slopes, then adding other
predictors as appropriate). Plot the data and fitted model to make sure that
your model makes sense.

(b) Use Bugs to simulate replicated datasets from your model, and make various
plots to compare the replicated with the actual data.


```{r q14_1}
library(rjags)
library(R2jags)
library(arm)
library(coda)
dogs <- read.table("dogs2.txt", row.names=1)


the.model <- "model{
for (j in 1:30){
        n.avoid[j,1] <- 0
        n.shock[j,1] <- 0
        for (t in 2:25){
            n.avoid[j,t] <- n.avoid[j,t-1] + (1 - y[j,t-1])
            n.shock[j,t] <- n.shock[j,t-1] + y[j,t-1]
        }
        for (t in 1:25){
            y[j,t] ~ dbin (p[j,t], 1)
            logit(p[j,t]) <- b.0 + b.1*n.avoid[j,t] + b.2*n.shock[j,t]
        }
    }
    b.0 ~ dnorm (0,0.01)
    b.1 ~ dnorm (0,0.01)
    b.2 ~ dnorm (0,0.01)
}"


# Defining the data you will pass into the model -- you *already know* these values
y <- dogs
dogs.data <- list ("y")

# Defining the initial values that your model's parameters (values you *don't* already know)
dogs.inits <- function (){list(b.0 = rnorm(1), 
                               b.1 = rnorm(1),
                               b.2 = rnorm(1))}
# Parameters to be returned
dogs.parameters <- c ("b.0", "b.1", "b.2")

# Now, we can actually run the model with the jags() function
dogs.1 <- jags( data=dogs.data, 
                inits=dogs.inits, 
                parameters.to.save=dogs.parameters, 
                model.file=textConnection(the.model), # Note the textConnection() function
                n.chains=3, 
                n.iter=5000, 
                DIC=F)
dogs.1


n.dogs <- nrow(dogs); n.trials <- ncol(dogs)
b.0 <- 1.81; b.1 <- -0.38; b.2 <- -0.21
n.sims <- 1000
y.rep <- array (NA, c(n.sims, n.dogs, n.trials))
for (j in 1:n.dogs){
    n.avoid.rep <- rep (0, n.sims)
    n.shock.rep <- rep (0, n.sims)
    for (t in 1:n.trials){
        p.rep <- invlogit (b.0 + b.1*n.avoid.rep + b.2*n.shock.rep)
        y.rep[,j,t] <- rbinom (n.sims, 1, p.rep)
        n.avoid.rep <- n.avoid.rep + 1 - y.rep[,j,t]
        n.shock.rep <- n.shock.rep + y.rep[,j,t]
    }
}

#Plot the results
par(mfrow=c(1,2),mar=c(2,2,2,2),oma=c(2,2,4,2),col.axis="white",col.main="white",
    col.lab="white",col.sub="white",col="white",bg="slategray")
plot(1:25,1-apply(dogs,2,mean),type="n")
rand.sample <- sample(x=1:n.sims,size=20)
for (i in rand.sample) lines(1:25,1-apply(y.rep[i,,],2,mean),col="mintcream")
lines(1:25,1-apply(dogs,2,mean),col="darkorchid4",lwd=3,type="l")
mtext(side=3,line=1.25,"T(yrep) versus T(y)")
plot(1:25,seq(-0.5,0.5,length=25),type="n")
for (i in rand.sample) 
    lines(1:25,(1-apply(y.rep[i,,],2,mean))-(1-apply(dogs,2,mean)),
    col="mintcream",lwd=2,type="l")
abline(h=0,col="darkorchid4",lwd=3)
mtext(side=3,line=1.25,"T(yrep)−T(y)")
title(outer=TRUE,line=2,col="white","Avoidance By Trial Number")


```