---
title: "HW6"
author: "Avinash Ramu"
date: "October 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prelude, include=FALSE}
library(ggplot2)
library(lme4)
library(arm)
```

##16.1
Elements of a Bugs model: list the elements of the model on page 370 by category:
modeled data, unmodeled data, modeled parameters, unmodeled parameters,
derived quantities, and looping indexes (as in Figure 16.4).

```
model {
  for (i in 1:n){
    y[i] ~ dnorm (y.hat[i], tau.y)
    y.hat[i] <- a[school[i]] + b*x[i]
  }
  b ~ dnorm (0, .0001)
  tau.y <- pow(sigma.y, -2)
  sigma.y ~ dunif (0, 100)
  for (j in 1:J){
    a[j] ~ dnorm (a.hat[j], tau.a)
    a.hat[j] <- g.0 + g.1*T[j]
  }
  g.0 ~ dnorm (0, .0001)
  g.1 ~ dnorm (0, .0001)
  tau.a <- pow(sigma.a, -2)
  sigma.a ~ dunif (0, 100)
}
```
- modeled data - y 
- unmodeled data - x, T
- modeled parameters - a
- unmodeled parameters - b, sigma.y, g.0, g.1, sigma.a
- derived quantities - y.hat, tau.y, a.hat, tau.a 
- looping indexes - i,j

##16.2
I have corrected the errors in the code below,
```r
model { #Bugs code
  for (i in 1:n){
    y[i] ~ dnorm (yhat[i], tau.y)
    yhat[i] <- a[state[i]] + theta*treat[i] + b*hispanic
  }
  theta ~ dnorm (0, .0001)
  b ~ dnorm (0, 1000)
  tau.y <- pow (sigma.y, -2)
  sigma.y <- dunif (0, 100)
  for (j in 1:J){
    a[j] ~ rnorm (mu.a, tau.a^2)
  }
  mu.a ~ dnorm (0, .0001)
  tau.a <- pow (sigma.a, -2)
  sigma.a ~ dunif (0, 100)
}
```