---
title: "Multilevel Modeling"
subtitle: 'Answer Key: Homework 1'
author: "Andy Stone"
output: pdf_document
header-includes:
 \usepackage{dcolumn}
 \usepackage{subcaption}
 \usepackage{adjustbox}
 \usepackage{graphicx}
---

\section{2.2}
\subsection{a.} The actual standard deviation is found by: 

$$s = \sqrt{\frac{\sum_{i=1}^{N} (x_{i} - \bar{x})^{2}}{N-1}} = 0.0064$$

The expected standard deviation is found by: 

$$sd^{expected} = \sqrt{\hat{p}(1-\hat{p})/n} = 0.008$$

```{r, echo=TRUE, collapse=TRUE, comment=NA, warning=FALSE, message=FALSE}
births <- c(.4777, .4875, .4859, .4754, .4874, .4864, .4813, .4787, .4895, .4797, .4876, 
            .4859, .4857, .4907, .5010, .4903, .4860, .4911, .4871, .4725, .4822, .4870, 
            .4823, .4973)

# Standard deviation of the proportions
sd(births)

# Expected standard deviation
(expected <- sqrt(mean(births) * (1 - mean(births)) * (1/3900)))
```


\subsection{b.} To determine if the expected and actual standard deviations are statistically distinguishable, we can use a two-sided chi-square test. This has the associated test statistic: 

$$ T = \frac{(N-1) * s^2}{\sigma^{2}} $$

```{r, echo=TRUE, collapse=TRUE, comment=NA, warning=FALSE, message=FALSE}
# Test statistic
print(chi.squared.statistic <- (23 * (0.0064)^2)/(0.008^2))

# Critical values
c(qchisq(0.025, 23), qchisq(0.975, 23))
```

The test statistic is contained within the critical values, so the difference is not statistically significant.


\section{2.3} The normal density curve is very similar to the histogram. This is to be expected, as a result of the Central Limit Theorem. The CLT says that, no matter the underlying distribution, the sum of many independent random variables (from a well-behaved distribution) will be an approximately normally distributed random variable.

```{r, echo=T, collapse=TRUE, comment=NA, warning=FALSE, message=FALSE, fig.height=3.8, fig.width=6}
set.seed(100) # For replication
xvals <- NULL # Empty vector to add simulated values of x into

# Creating the 1000 simulations of x
for(i in 1:1000){
  xvals <- c(xvals, sum(runif(20, min=0, max=1)))
}

par(mfrow=c(1,1)) # Plotting parameters
h <- hist(xvals, plot=FALSE) # Characteristics of the histogram
ylim1 <- range(0, h$density, dnorm(0)) # Setting range of y values for histogram

# Plotting histogram
hist(xvals, freq=FALSE, ylim=ylim1, main="Figure 1: Histogram of Simulated x Values
     and Normal Density Curve", xlab="x Values", cex.lab=0.7, cex.main=0.9, cex.axis=0.7)
# Overplotting normal density
curve(dnorm(x, mean=mean(xvals), sd=sd(xvals)), add=TRUE, col="red")
```







