---
title: "HW5"
author: "Avinash Ramu"
date: "October 11, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prelude, include=FALSE}
library(ggplot2)
library(lme4)
library(arm)
library(foreign)
```

## Q14.5
###A

$$Pr(dec = 1) = logit^{-1}(\beta_0 + \beta_1 attr + \beta_2 sinc + \beta_3 intel + \beta_4 fun + \beta_5 amb + \beta_6 shar)$$

```{r q1_1}
sd <- read.csv("speed_dating.csv", head = T)
sd[is.na(sd)] <- 0
attach(sd)
fit.1 <- glm (dec ~ attr + sinc + intel + fun + amb + shar, family=binomial(link="logit"))
summary(fit.1)
```

Attr, fun and shar have a positive coefficient and increase the chance of a repeat-date. The coefficients are significant. sinc and amb also have significant coefficients and reduce the chance of a repeat date.

A 1 point increase in attractiveness increases the probability of a repeat date by 14%
A 1 point increase in fun increases probability of repeat date by about 5.5%
A 1 point increase in shar increases probability of a repeat date by about 3%
A 1 point increase in sinc reduces probability of a repeat date by about 1%
A 1 point increase in amb reduces probability of a repeat date  by about 2.7%

###B
Add a explanatory variable `iid`

$$Pr(dec = 1) = logit^{-1}(\beta_0 + \beta_1 attr + \beta_2 sinc + \beta_3 intel + \beta_4 fun + \beta_5 amb + \beta_6 shar + \beta_7 iid)$$

```{r q1_2}
fit.2 <- glm (dec ~ attr + sinc + intel + fun + amb + shar + iid, family=binomial(link="logit"))
summary(fit.2)
```
The coefficients are similar to the previous model. The rater does not have a significant coefficient.

###C
Add a explanatory variable `pid`
$$Pr(dec = 1) = logit^{-1}(\beta_0 + \beta_1 attr + \beta_2 sinc + \beta_3 intel + \beta_4 fun + \beta_5 amb + \beta_6 shar + \beta_7 iid + \beta_8 pid)$$

```{r q1_3}
fit.3 <- glm (dec ~ attr + sinc + intel + fun + amb + shar + iid + pid, family=binomial(link="logit"))
summary(fit.3)
```

Both the rater and the person being rated now have significant coefficients. This indicates that certain raters have
preferences for certain kinds of dates.

## Q14.6


###A

I used the average of the six predictors as a explanatory variable. I set the NA's to zero.
$$Pr(dec_i = 1) = logit^{-1}(\beta_{i0} + \beta_{i1} newscore)$$
```{r q2_1}
sd$newscore <- (sd$attr + sd$sinc +  sd$intel + sd$fun + sd$amb +  sd$shar)/6
sd$iid <- as.factor(sd$iid)
some_iids <- sample(sd$iid, 5)
for (iid1 in some_iids) {
  sd_subset <- sd[sd$iid == iid1, ]
  fit.21 <- glm (sd_subset$dec ~ sd_subset$newscore, family=binomial(link="logit"))
  print(summary(fit.21))
}

```


The average predictor does not come out to be significant for either of the raters, we appear
to lose information by averaging over all the predictors.

###B
Vary the intercept and coefficients by group of the rater.
$$Pr(dec = 1) = logit^{-1}(\beta_{j0} + \beta_{j1} attr + \beta_{j2} sinc + \beta_{j3} intel + \beta_{j4} fun + \beta_{j5} amb + \beta_{j6} shar + \beta_{j7} iid + \beta_{j8} pid)$$

```{r q2_2}
fit.lmer1 <- lmer(dec ~ 1 + (1 + attr + sinc + intel + fun + amb + shar | iid), family=binomial(link="logit"), data=sd)
display(fit.lmer1)
```

Attractiveness seems to vary by the rater the most. Other significant predictors that are helpful at the group level are sinc, intel, fun and share. Amb could probably be removed from the grouping based on the SD explained. The raters also have significant intercepts indicating that grouping by rater is a good idea.

###C

I have included comments under each model. The no-pooling model cannot be interpreted since the average score does not seem to be a  significant predictor. The residual deviance is lowest under the model in 14.6 (B), the multilevel model grouped by the rater. The residual deviance in the model in 14.5 (A) is twice as much as the multilevel model. Grouping by rater seems to give us the best model in this example.

##15.1
$$presvote\_intent = \beta_0 + \beta_1 gender + \beta_2 race + \beta_3 age + \beta_4 educ1 + \beta_5 income + \beta_6 religion + \alpha_{j[i]}$$

##15.2
###A
$$presvote\_intent = \beta_0 + \beta_1 gender + \beta_2 race + \beta_3 age + \beta_4 educ1 + \beta_5 income + \beta_6 religion + \alpha_{j[i]} + \beta_{7j[i]} ideology$$

###B
```{r q4_2}
nes <- read.dta("nes5200_processed_voters_realideo.dta")
attach(nes)
nes$presvote_intent <- as.numeric(nes$presvote_intent)
fit.lmer1 <- lmer( presvote_intent ~ 1 + gender + race + age + educ1 + income + religion + (1 + real_ideo | state), data = nes)
summary(fit.lmer1)
```

The model failed to converge. The real_ideo variable does not seem to show a strong effect of grouping by state. The SD explained by this variable at the group level is a fraction of the residual SD.
