---
title: "HW3"
author: "Avinash Ramu"
date: "September 25, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r prelude, include=FALSE}
library(ggplot2)
library(lme4)
library(arm)
```

## Q11.4
###A 
Let's use a random sample of 10 patients to plot since 250 is a large number to display
```{r q1_1}

cd4 <- read.table("dat/allvar.csv", head = T, sep = ",")
cd4$time <- cd4$visage - cd4$baseage
cd4_subset <- cd4[cd4$newpid %in% sample(cd4$newpid, 10), ]
cd4$newpid <- as.factor(cd4$newpid)
(plot1 <- ggplot(cd4_subset,aes(x=time, y=sqrt(CD4PCT))) + geom_point() + facet_wrap(~newpid))
```

###B
Print the models fit to the ten patients
```{r q1_2}
cd4_splits <- split(cd4_subset, cd4_subset$newpid)
for (i in 1:length(cd4_splits)) {
  lm1 = lm(sqrt(CD4PCT) ~ time, data = cd4_splits[[i]])
  coeffs <- lm1$coeff
  plot1 <- ggplot(data = cd4_splits[[i]]) + geom_abline(slope = coeffs[2], intercept = coeffs[1]) + geom_point(aes(x = time, y = sqrt(CD4PCT)))
  print(plot1)
}

```

###C
Setup a model in which the treatment and age affect outcome
```{r q1_3}

lm1 = lm(sqrt(CD4PCT) ~ treatmnt + time + factor(newpid) - 1, data = cd4)
#print(summary(lm1))
```
The coefficient for patient 31 is 0.70309, we can fit this to that patients data
and set this coefficient as the outcome variable to get group level alphas and betas.
```{r q1_3_2}
patient_31 <- cd4[cd4$newpid == 31, ]
```


## Q12.2
###A
```{r q2_1}
lmer1 <- lmer(sqrt(CD4PCT) ~ time + (1|newpid), data=cd4)
print(summary(lmer1))
```
The coefficient for time is -0.36609 with a SE of 0.05399, the coefficient is less than
zero in the +/- 2 * SE interval hence significant. This indicates a decrease in CD4PCT
with time. The intercepts vary by patient.


###B
```{r q2_2}
lmer2 <- lmer(sqrt(CD4PCT) ~ time + (treatmnt + baseage | newpid), data=cd4)
summary(lmer2)
```
The coefficient for time is still the same as part A and indicates a decrease in CD4PCT with time. Note - The model fails to converge. The variance of the treatment and baseage coefficients for the different groups is close to zero. This is indicates that the treatment and baseage have a similar effect across groups.

###C
```{r q2_2_2}
intercepts <- fixef(lmer2)[1] + ranef(lmer2)$newpid[,1]
slope1 <- ranef(lmer2)$newpid[,2] #slope for treatmnt
slope2 <- ranef(lmer2)$newpid[,3] #slope for baseage

```


## Q12.5
```{r q3_1}
radon <- read.table("dat/mn-radon-2.txt")
head(radon)
lmer1 <- lmer(log.radon ~ 1 + (1 + county_sample_size | county), data = radon)
summary(lmer1)
#coef(lmer1)
```
The county sample size has a very small sigma compared to the residual sigma, this grouping
doesn't seem to help the model a lot.
